---
title: "Exploratory analysis"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(ggseas)
```


```{r data}
calendar <- read_csv("Data/calendar.csv")
sales_train_validation <- read_csv("Data/sales_train_validation.csv")
sell_prices <- read_csv("Data/sell_prices.csv")
```


## Plan
We are interested in predicting the sales of TX_1. Therefore we will mainly focus on this store.

calendar:
- List of event_type_1 and event_type_2
- Barplot. X axis = weekdays, Y axis = count(event)
- Barplot. X axis = event_type, Y axis = count(event)
- What are the events for which snap_TX = 1?
- Time series. Y axis = count(event). Daily? Weekly? Monthly? Yearly? (will be usefull to compute correlation with time series about sales)
- Same as precedent but only when snap_TX = 1
- Résumer dans une table le nombre d'events par semaine, mois, année pour chaque année

sell_prices:
- Min, Max, Mean price for TX_1
- Compared to the ohter stores?

sales_train_validation:
- First we need to join the sell prices
- Sales per product (#, $ or %) for TX_1
- Sales per category (#, $ or %) for TX_1
- Total sales (# and $) for TX_1
- Barplot. X axis = c(all, TX, TX_1), Y axis = % of sales
- Time series of sales for TX_1
* autoplot()
* ggseasonplot()
* ggsubseriesplot()
* gglagplot()
* ggacf()
* checkresiduals()
- qplot using ts of sales and ts of events? (any correlation?)
